A Deeper Look into Bug Fixes: Patterns, Replacements, Deletions, and Additions

### 2016-03-11 MSRChallenge2016.tex
boom.  4 pages.

There has been considerable attention paid to automatic
program repair in recent
years (e.g.,~\cite{kim2013,legoues2012,
  Mechtaev15,Long2016}).
%
One broad class of techniques in this space take a generate-and-validate
approach: \emph{generating} a large number of candidate patches using a
pre-defined set of mutation operators, and then \emph{validating} correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

@@

\vspace{1ex} \noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} ask
some of the same questions we do, on  5 projects. We study a much
larger dataset~\cite{dyer2013} with 380,125 repositories; and we look at more
statement types, as well as replacements. Similar to our study in
Section~\ref{sec:stmtstudy}, researchers have 
studied AST-level~\cite{Martinez:2015ez} and line-level
changes~\cite{Asaduzzaman:2013gu, Asaduzzaman:2013df} across bug fixing commits.
Although the granularity differs, our approach is also novel with respect to scale of
analysis. Barr et al.~\cite{Barr14fse} studied changes to Java programs to
understand the ``Plastic Surgery Hypothesis'' underlying certain types of
program repair, an orthogonal concern. Kim et al.~\cite{kim2013} manually
analyze changes to Java programs to inform automated repair, and
show that doing so results in higher-quality repairs; we study their templates
heuristically on a different dataset. Their results motivate studies of human
repairs, as it may result in better patches.  Long and Rinard learn
probabilistic models from bug fixes to C code~\cite{Long2016}. Our analysis is
significantly less precise and does not inform a new technique.  Rather, it
serves as a starting point for research on the repair of Java bugs.

@@

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
\bottomrule
  \toprule
  Introducing & Total Number & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.2 \\ 
\bottomrule
  \end{tabular}
  \caption{Top: File Types of Changed Files.  Bottom: Introduction of new
    classes, methods, fields, local variables to fix bugs.   \label{tbl:fileType}\label{tbl:new}}
\vspace{-0.5cm}
\end{table}
\vspace{1ex}
\noindent\textbf{What are the types of those changed files?}
%
The top of Table~\ref{tbl:fileType} shows the types of the changed files in the
Boa dataset. 
Text and
binary files are changed least frequently. This is unsurprising, since such
files are often documentation, and binaries should be changed rarely. 
%
XML files in Java projects usually represent build files;
changing build files is unsurprising.  Rather more surprising is how frequently
UNKNOWN files are changed. Given that all projects are Java projects, it is
unclear whether these are written in other programming languages. We
hypothesize they might be test resources.  JAVA (ERROR) files are those that Boa failed to parse. JLS2, JLS3, and JLS4 are
Java files in different J2SE versions.  If we group
all Java types, on average, each bug fixing
revision changes $4.47$ Java files. This does not conform to our assumption of
bug locality.  To better understand what changes
are usually made to those Java files, we next ask:

@@

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This template changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre- and post-fix versions
of a buggy file, we create a custom method call signature that
contains the name string, literal parameters string, and variable parameters string (complex
expressions are listed as ``OTHER'' string).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre- and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different signatures between versions.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
template changes a method name. % (\texttt{obj.method1(param) $\rightarrow$
%obj.method2(param)}).  
To detect it, we create a method signature similar
to the above for AMP, looking for methods
with the exact same parameter signature but different names in pre- and
post-fix versions. %If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
template adds a parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create a method signature similar to the above, seeking methods with
  the same name, but with one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Add or remove a branch condition (ABC)}. This template adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
We count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre- and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a count difference between pre- and post-fix version,
  we consider an instance of the ABC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This template adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that an instance of the IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This template inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre- and post-fix
  version. If there is a larger count in post-fix version,
  we consider that an instance of the ANC pattern found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
%	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
 We count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre- and
  post-fix version. If there is a larger count in post-fix version, we consider that an instance of the AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre- and
post-fix version. If the is a larger count in post-fix
version, we consider that an instance of the COB pattern is found.

\vspace{1ex} The above patterns cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are (1) Class Cast Checker (cannot be analyzed because
Boa does not support the {\em instanceof} expression); and (2) Expression Changer (requires us to track scope between pre- and post-fix version; not
easily supported by Boa). Investigating 8 out of
10 patterns still shows us how common the patterns actually are.

@@

Some automatic program repair approaches seek generality by using
higher-granularity mutation operators such as statement-level addition, deletion and replacement. To
support the generation of high-quality patches, we 
next analyze how developers mutate source code to fix bugs at this granularity
level.

\subsection{Detection Approach}
Because direct diffs are difficult to identify on this dataset, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another.
For each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

We follow a similar approach to approximate a count of deletions and
insertions. 
For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

@@

\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the Do statement row (row 5) and Assert statement column (column 2)
shows 0.81, it means that Do statements were
replaced by Assert statements 0.81\% of the times.

Additional results (raw numbers not shown) show that the most common replacement
is replacing Return statement with If statement (appearing in 30,489 files). The
second most common replacement is replacing an If statement with a Return
statement, which had 28,536 appearances. 
By contrast, the least common replacement was an Assert
statement replacing a Type Declaration statement, which we did not observe. 
The second least common replacement was a Do statement for a Type
Declaration or a Label with a Type; we observed these once each.

The most common
\emph{replacer} statement kind (the statement kind that most commonly replaces
others) is the \texttt{If Statement} (101,366 appearances). The least
common \emph{replacer} is the \texttt{Type Declaration Statement} with 447
appearances.
The most common \emph{replacee} was the
\texttt{Return Statement} with 111,938 appearances; and the least common
\emph{replacee} was the \texttt{Type Declaration}, with 399 appearances.

\vspace{1ex}
\noindent \textbf{Potential deletions and insertions} From Figure~\ref{fig:StmtType}, we
can see that, expression, if, return, for and try catch statements are added
most often. Meanwhile, the same types of statements are also deleted most
often. This finding indicates that most bugs were fixed by changing the control
flow of the program.

An important study that complements for our study is the repair model approach
proposed by Martinez and Monperrus~\cite{Martinez:2015ez}, in which they discuss
a probability distribution for when to apply which kind of edit. Although their
approach is able to trace more fine-grain AST-level changes, our results are
consistent with their findings. For example, their empirical
analysis~\cite{Martinez:2015ez} shows that method invocations, if statements,
and variable declaration statements are added/deleted/updated most often, which
is also illustrated in Figure~\ref{fig:StmtType} (In Boa, method invocation and
variable declaration fall into the category of expression). It is also worth
mentioning that their study covers 14 repositories with 89,993 versioning
transactions, while our study studies 380,125 repositories with 23,229,406
revisions.

@@

\noindent\textbf{Threats to validity.} Since we are using a DSL
provided by Boa, the correctness of our analysis
depends on both our programs and Boa itself. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem.  To mitigate the risk of errors in
implementation, we open source our scripts\footnote{\url{https://github.com/chupanw/BoaChallenge}}.
Because Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions, our template
matching and analysis of code changes (by counting each statement kind) only
provide estimates of behavior; we consider our results as informative
approximations.

>>>>

There has been considerable recent attention paid to automatic
program repair (e.g.,~\cite{kim2013,legoues2012,
  Mechtaev15,Long2016}).
%
One broad class of techniques in this space take a generate-and-validate
approach: \emph{generating} a large number of candidate patches using a
pre-defined set of mutation operators, and then \emph{validating} correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

@@

\vspace{1ex} \noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} ask
some of the same questions we do, on  5 projects. We study a much
larger dataset~\cite{dyer2013} with 380,125 repositories; and we look at more
statement types, as well as replacements. Similar to our study in
Section~\ref{sec:stmtstudy}, researchers have 
studied AST-level~\cite{Martinez:2015ez} and line-level
changes~\cite{Asaduzzaman:2013gu} across bug fixing commits.
Although the granularity differs, our approach is also novel with respect to scale of
analysis. Barr et al.~\cite{Barr14fse} studied changes to Java programs to
understand the ``Plastic Surgery Hypothesis'' underlying certain types of
program repair, an orthogonal concern. Kim et al.~\cite{kim2013} manually
analyze changes to Java programs to inform automated repair, and
show that doing so results in higher-quality repairs; we study their templates
heuristically on a different dataset. Their results motivate studies of human
repairs, as it may result in better patches.  Long and Rinard learn
probabilistic models from bug fixes to C code~\cite{Long2016}. Our analysis is
significantly less precise and does not inform a new technique.  Rather, it
serves as a starting point for research on the repair of Java bugs.

@@

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
\bottomrule
  \toprule
  Introducing & Total & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.20 \\ 
\bottomrule
  \end{tabular}
  \caption{Top: File Types of Changed Files.  Bottom: Introduction of new
    classes, methods, fields, local variables to fix bugs.   \label{tbl:fileType}\label{tbl:new}}
\vspace{-0.5cm}
\end{table}
\vspace{1ex}
\noindent\textbf{What are the types of those changed files?}
%
The top of Table~\ref{tbl:fileType} shows the types of the changed files in the
Boa dataset. 
Text and
binary files are changed least frequently. This is unsurprising, since such
files are often documentation, and binaries should be changed rarely. 
%
XML files in Java projects usually represent build files;
changing build files is unsurprising.  Rather more surprising is how frequently
UNKNOWN files are changed. Given that all projects are Java projects, it is
unclear whether these are written in other programming languages. We
hypothesize they might be test resources.  JAVA (ERROR) files are those that Boa failed to parse. JLS2, JLS3, and JLS4 are
Java files in different J2SE versions.  If we group
all Java types, on average, each bug fixing
revision changes $4.47$ Java files. This does not conform to our assumption of
bug locality.  To better understand what changes
are usually made to those Java files, we next ask:

@@

%\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This template changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre- and post-fix versions
of a buggy file, we create a custom method call signature that
contains the name, literal parameters, and variable parameters, all as string (complex
expressions are listed as ``OTHER'' string).
We discard signatures that appear both in pre- and post-fix
versions, and then identify methods with the same name and 
number of parameters, but different signatures between versions.

%\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
template changes a method name. % (\texttt{obj.method1(param) $\rightarrow$
%obj.method2(param)}).  
To detect it, we create a method signature similar
to the above for AMP, looking for methods
with the exact same parameter signature but different names in pre- and
post-fix versions. %If it exists, we consider that MSM pattern is found.

%\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
template adds a parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create a method signature similar to the above, seeking methods with
  the same name, but with one more parameter in the post-fix version.

%\vspace{1ex}
\noindent
 {\bf Add or remove a branch condition (ABC)}. This template adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
We count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre- and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a difference in count between versions,
  we consider an instance of the ABC pattern is found.

%\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This template adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that an instance of the IAO pattern is found.

%\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This template inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre- and post-fix
  version. If there is a larger count in post-fix version,
  we consider that an instance of the ANC pattern found.

%\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
%	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
 We count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre- and
  post-fix version. If there is a larger count in post-fix version, we consider that an instance of the AOB pattern is found.

%\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre- and
post-fix version. If the is a larger count in post-fix
version, we consider that an instance of the COB pattern is found.

These patterns cover 8 of the 10 bug fixing patterns in
PAR. We do not investigate (1) Class Cast Checker (cannot be analyzed because
Boa does not support the {\em instanceof} expression); and (2) Expression Changer (requires us to track scope between pre- and post-fix version; not
easily supported by Boa). Investigating 8 out of
10 patterns provides an informative approximation of their prevalence.

@@

Some automatic repair approaches seek generality by using
higher-granularity mutation operators such as statement-level addition, deletion and replacement. To
support the generation of high-quality patches, we 
analyze how developers mutate source code to fix bugs at this granularity
level.

\subsection{Detection Approach}
Because direct diffs are difficult to identify on this dataset, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another.
For each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.
%
We follow a similar approach to approximately count deletions and
insertions. 
For each bug fixing revision $r$ and each statement kind $k$, we compare the
count of statements of kind $k$ in revision $r$ and $r-1$.

@@

\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, 
the corresponding to the $\mathtt{Do}$ row (row 5) and $\mathtt{Assert}$ column (column 2)
shows 0.81, indicating that $\mathtt{Do}$ statements were
replaced by $\mathtt{Assert}$ statements 0.81\% of the times.

Additional analysis (raw numbers not shown) show that the most common replacement
replaces $\mathtt{Return}$ statements with $\mathtt{If}$ statements (in 30,489 files). The
second most common replacement replaces an $\mathtt{If}$ statement with a $\mathtt{Return}$
(28,536 incidences). 
By contrast, the least common replacement was an $\mathtt{Assert}$
statement replacing a $\mathtt{TypeDecl}$, which we did not observe. 
The second least common replacements were replacing $\mathtt{Do}$ statements or $\mathtt{Label}$s for a $\mathtt{TypeDecl}$; we observed these once each.

The most common
\emph{replacement} statement (the statement kind that most commonly replaces
others) is the $\mathtt{If}$ statement (101,366 appearances). The least
common \emph{replacement} is the $\mathtt{TypeDecl}$ (447
appearances).
The most common \emph{replacee} was the
$\mathtt{Return}$ (111,938 appearances); the least common
\emph{replacee} was again the $\mathtt{TypeDecl}$ (399 appearances).

\vspace{1ex}
\noindent \textbf{Potential deletions and insertions} From Figure~\ref{fig:StmtType}, we
can see that $\mathtt{expression}$, $\mathtt{If}$, $\mathtt{Return}$,
$\mathtt{for}$ and $\mathtt{Try}$ statements are both added and deleted
most often as compared to the other statement times. 
These findings indicate that most bugs were fixed by changing control
flow.

An important study that complements ours is the repair model approach
proposed by Martinez and Monperrus~\cite{Martinez:2015ez}, which proposes
a probability distribution suggesting when to apply which kind of edit. Although their
approach is able to trace more fine-grain AST-level changes, our results are
consistent with their findings. For example, their empirical
analysis~\cite{Martinez:2015ez} shows that method invocations, if statements,
and variable declarations are added/deleted/updated most often, which
is also illustrated in Figure~\ref{fig:StmtType} (Boa groups method invocations and
variable declarations into the $\mathtt{Expression}$ category).  Our study
complements there at a much larger scale (we study 380,125 repositories with 23,229,406
revisions as compared to the 14 repositories in the prior work).

@@

\noindent\textbf{Threats to validity.} The correctness of our analysis
depends on both our programs and Boa and its DSL. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem.  To mitigate the risk of implementation errors,
we released our scripts\footnote{\url{https://github.com/chupanw/BoaChallenge}}.
Because Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions, our template
matching and analysis of code changes (by counting each statement kind) only
provide estimates of behavior; we consider our results as informative
approximations.
